The Importance of Software and Computing to Particle Physics

### 2018-12-18 HSF-Notes/2018-01/espp-hsf-submission.tex
Final small updates

%% Comment out before final submission
\usepackage{lineno}  % for line numbering during review
\linenumbers

@@

\author{A contribution from the High-Energy Physics Software Foundation
to the European Strategy for Particle Physics}

@@

The current success of the experimental High-Energy Physics programme
relies hugely on software and computing. We have developed many
innovative and unique pieces of software as a necessity, as no
off-the-shelf solutions were available. The HEP software landscape
consists of tens of millions of lines of code, from widely used community
supported packages, like ROOT and Geant4, to experiment and detector
specific codes. Thousands of physicists spend time writing this code and
the lifetimes of HEP experiments require maintaining and evolving these
codes over decades. These codes are run at a massive scale. For the LHC
experiments, in particular, about 750k CPU cores are constantly employed
on the Worldwide LHC Grid running event generation, simulation,
reconstruction and analysis jobs that are critical to our physics
output. The costs of this part of our programme is very high - to
purchase just this CPU resource in a commercial cloud would cost hundreds of
millions of Euros every year. It is incumbent on us to use these
resources as efficiently as possible for physics.

>>>>

%% Comment out before final submission
%\usepackage{lineno}  % for line numbering during review
%\linenumbers

@@

\author{A contribution from the High-Energy Physics Software Foundation
to the European Particle Physics Strategy Update 2018-2020}

@@

The current success of the experimental High-Energy Physics programme
relies hugely on software and computing. We have developed many
innovative and unique pieces of software as a necessity, as no
off-the-shelf solutions were available. The HEP software landscape
consists of tens of millions of lines of code, from widely used community
supported packages, like ROOT and Geant4, to experiment and detector
specific codes. Thousands of physicists spend time writing this code and
the lifetimes of HEP experiments require maintaining and evolving this
software over decades. These codes are run at a massive scale. For the LHC
experiments, in particular, about 750k CPU cores are constantly employed
on the Worldwide LHC Grid running event generation, simulation,
reconstruction and analysis jobs that are critical to our physics
output. The cost of this part of our programme is very high - to
purchase just this CPU resource in a commercial cloud would cost hundreds of
millions of Euros every year. It is incumbent on us to use these
resources as efficiently as possible for physics.
